### 2025-08-17 correlation.qmd
20250816 - tweaks
20250817 - correlation ratio

You can compute a correlation manually using the *z*-scores, as in @eq-pearsonCorrelation:

@@

An example of a nonlinear correlation is a distance correlation and the xi ($\xi$) coefficient.
These estimates of nonlinear correlation only range from 0–1 (where 0 represents no association and 1 represents a strong association), so they are interpreted differently than a traditional correlation coefficient (which can be negative).
We can estimate a distance correlation using the `correlation::correlation()` function of the correlation package [@R-correlation; @Makowski2020_packages].

```{r}
correlation(
  data = player_stats_seasonal %>% select(draftround, fantasyPoints),
  method = "distance"
)
```

@@

```{r}
correlation(
  data = data.frame(v1, v2),
  method = "biweight"
)

correlation(
  data = data.frame(v1, v2),
  method = "percentage"
)
```

>>>>

You can compute a correlation manually using the [*z* scores](#sec-zScores), as in @eq-pearsonCorrelation:

@@

Examples of a nonlinear correlation include the correlation ratio [$\eta$; @Goodwin2006], distance correlation, and the xi ($\xi$) coefficient [@Chatterjee2021].
These estimates of nonlinear correlation only range from 0–1 (where 0 represents no association and 1 represents a strong association), so they are interpreted differently than a traditional correlation coefficient (which can be negative).

Here is how to compute the correlation ratio (eta; $\eta$) using the `stats::aov()` and `effectsize::eta_squared()` functions:

```{r}
mod <- stats::aov(
  fantasyPoints ~ height,
  data = player_stats_seasonal)

correlationRatio <- effectsize::eta_squared(mod)

cols_to_sqrt <- intersect(
  colnames(correlationRatio),
  c("Eta2", "CI_low", "CI_high"))

correlationRatio[cols_to_sqrt] <- sqrt(correlationRatio[cols_to_sqrt])

correlationRatio <- correlationRatio %>%
  rename(Eta = Eta2)

correlationRatio
```

We can estimate a distance correlation using the `correlation::correlation()` function of the correlation package [@R-correlation; @Makowski2020_packages].

```{r}
correlation::correlation(
  data = player_stats_seasonal %>% select(height, fantasyPoints),
  method = "distance"
)
```

@@

```{r}
correlation::correlation(
  data = data.frame(v1, v2),
  method = "biweight"
)

correlation::correlation(
  data = data.frame(v1, v2),
  method = "percentage"
)
```
